### 2020-10-06 latex/vlp.tex
Fully edited the intro to VLP

The motion of the VLP has been well studied (see for instance
\cite{dynamics_periodic_vlp}), and many control mechanisms exist
to stabilize trajectories of the system. While many of these controllers 
are time-dependent, \citet{vlp_energy_shaping}
offer a time-independent technique to inject energy into the VLP. 
They design a controller through a technique called \textit{energy shaping},
and prov that their control mechanism stabilizes any desired energy level set.
However, the energy injection mechanism is ad-hoc in the sense that it is not
derived from human behaviour; their technique is tailored to this specific
problem, and cannot be generalized easily.
In this chapter we will make use of the general VNHC framework developed in Chapter
\ref{ch:vnhcs} to stabilize energy levels of the VLP in a
time-independent manner. 
These VNHCs will add and remove energy from the VLP while maintaining the
structured motion of a human on a swing.

\section{Dynamics of the Variable Length Pendulum}
% Derive the dynamics of the VLP in Hamiltonian form
We will model the VLP as a point mass \(m\)
connected to a fixed pivot by a massless rod of varying length \(l\) with angle 
\(q \in \mathbb{S}^1\) from the vertical, as is seen in Figure
\ref{fig:vlp-model}. 
We will also ignore any damping and frictional forces in this model.
In a realistic VLP, the rod length \(l\) varies between some minimum
length \(\underline{l} \geq 0\) and some maximum length 
\(\overline{l} > \underline{l}\). The configuration of the VLP is the vector
\(\mathbf{q} := (q,l) \in \mathbb{S}^1 \times [\underline{l},\overline{l}]\).

@@

Using this configuration, we will compute the Hamiltonian dynamics of the system.
The cartesian position of the mass at the tip of the pendulum
is given by \(x = (l\sin(q),-l\cos(q))\), while its velocity is
\(\dot{x} = (\dot{l}\sin(q) + l\cos(q)\dot{q}, -\dot{l}\cos(q) + l\sin(q)\dot{q})\).
Computing the kinetic energy \(T\) yields
\[
   T(\mathbf{q},\dot{\mathbf{q}}) = 
   \frac{1}{2}m\norm{\dot{x}}^2 = \frac{1}{2}m\left(\dot{l}^2 + l^2\dot{q}^2\right)
   .
\]
The potential energy \(P\) with respect to the pivot (under a gravitational
acceleration \(g\)) is
\[
   P(\mathbf{q}) = -mgl\cos(q)
   .
\]
Collecting the kinetic energy into a quadratic form, we get the Lagrangian
\[
   \mathcal{L}(\mathbf{q},\dot{\mathbf{q}}) 
   = \frac{1}{2} \dot{\mathbf{q}}\tpose D(\mathbf{q})\dot{\mathbf{q}} - P(\mathbf{q})
   = \frac{1}{2}
   \begin{bmatrix} \dot{q} & \dot{l} \end{bmatrix}
   \begin{bmatrix}
      ml^2 & 0 \\
      0 & m \\
   \end{bmatrix}
   \begin{bmatrix} 
      \dot{q} \\ \dot{l}
   \end{bmatrix}
   + mgl\cos(q)
   .
\]
Computing the conjugate of momenta to \(\mathbf{q}\), we get 
\[
   \mathbf{p} := \begin{bmatrix} p \\ p_l \end{bmatrix} 
   = \begin{bmatrix} ml^2\dot{q} \\ m\dot{l} \end{bmatrix} 
   .
\]
Performing the Legendre transform on \(\mathcal{L}\) and setting
\(M(\mathbf{q}) := D(\mathbf{q})\), \(V(\mathbf{q}) := P(\mathbf{q})\),
we get the Hamiltonian (\ref{eqn:simply-actuated-hamiltonian}).
Taking the appropriate derivatives, the dynamics of the VLP in Hamiltonian form
are described in (\ref{eqn:vlp-hamiltonian-with-pl}). 
\begin{align}\label{eqn:vlp-hamiltonian-with-pl}
   \mathcal{H} &= \frac{1}{2} \begin{bmatrix} p & p_l \end{bmatrix}
      \begin{bmatrix}
         \frac{1}{ml^2}  & 0 \\
         0 & \frac{1}{m}
      \end{bmatrix} \begin{bmatrix} p \\ p_l \end{bmatrix} - mgl\cos(q) \\
     &\begin{cases}
        \dot{q} = \frac{p}{ml^2} \\
        \dot{l} = \frac{p_l}{m} \\
        \dot{p} = -mgl\sin(q) \\
        \dot{p}_l = \frac{p^2}{ml^3} + mg\cos(q) + \tau \\
   \end{cases} \nonumber
\end{align}
The control input is a force \(\tau \in \mathbb{R}\) affecting the dynamics of
\(p_l\), acting colinearly with direction of the rod.
We assume the force does not affect the dynamics of \(p\) in any way -
that is, the control input cannot enact any lateral force on the pendulum.
This is extremely useful because it means the VLP is described in simply
actuated coordinates, which allows us to apply the theory of VNHCs we
developed in Chapter \ref{ch:vnhcs}.
% TODO: Replace this with a description that the dynamics of (q,p) are the
% constrained dynamics, and that we are using a VNHC o f the form h = l -
% L(q,p). We will simplify notation by saying that l = L(q,p) to mean that we
% are stabilizing h = l - L = 0
Let us define a ``VNHC" of the form \(l = L(q,p)\) (by this we mean we are
actually defining the VNHC \(h(q,p) = l - L(q,p) = 0\)).
Assuming this is a regular VNHC, the constraint manifold \(\Gamma\) is
\(\mathbb{S}^1 \times \R\), and it is parameterized by the unactuated phase
\((q,p)\).
The constrained dynamics are described entirely by
\((\dot{q},\dot{p})\) with \(l\) replaced by \(L(q,p)\) (omitting the inputs to
\(L\) for clarity):
\begin{equation}\label{eqn:vlp-dynamics}
   \begin{cases}
      \dot{q} = \frac{p}{m L^2} \\
      \dot{p} = -mgL\sin(q)
      . \\
   \end{cases}
\end{equation}

The total mechanical energy of the system restricted to the constraint manifold
is given by (\ref{eqn:vlp-energy}), where \(L(q,p)\) and \(\dot{L}(q,p)\) are
known.
\begin{equation}\label{eqn:vlp-energy}
   E(q,p) = \frac{p^2}{mL^2} + \frac{1}{2}\dot{L}^2 - mgL\cos(q)
\end{equation}
In the rest of this chapter, we will derive the function \(L(q,p)\) which
allows the constrained dynamics to inject energy into the VLP over time.

@@

\section{The VLP Constraint}
% Go through the development of the constraint and prove it gains energy
To motivate why a VNHC could inject energy into the VLP in a human-like
manner, we will examine a person standing on a swing.
As can be seen in Figure \ref{fig:child-vlp}, their center of mass moves closer
to the swing's pivot when they stand, and moves away from the pivot when they
squat.
This is equivalent to the VLP model from Figure \ref{fig:vlp-model}, where
standing and squatting correspond to shortening and lengthening the pendulum
(respectively).
\begin{figure}
   \centering
   \begin{subfigure}[t]{0.45\textwidth}
      \includegraphics[]{images/child_vlp_standing.png}
      \caption{A person standing on a swing has their center of mass 
      close to the pivot.}
   \end{subfigure}
   \hfill
   \begin{subfigure}[t]{0.45\textwidth}
      \includegraphics[]{images/child_vlp_squatting.png}
      \caption{When a person squats on a swing, their center of mass extends
      away from the pivot.}
   \end{subfigure}
   \caption{The VLP representation of a person on a standing swing.}
   \label{fig:child-vlp}
\end{figure}

@@

Notice that the child's pumping strategy requires knowledge of
when the system is at the ``bottom" or ``top" of the swing. 
Since being at the ``bottom" is equivalent to having angle \(q = 0\) and being
at the ``top" is equivalent to having momentum \(p = 0\), a controller based on
this strategy will necessarily use the full phase \((q,p)\). 
This is why we will use VNHCs instead of VHCs.

The time-optimal controller from \cite{pumping_swing_standing_squatting} is (in
our notation)
\[
   l_\text{opt}(q,p) := -\sign{qp}
\]
which is a piecewise-continuous controller that varies between \(\pm 1\). 
This looks similar to a VNHC, but it is not smooth and makes the assumption that 
\(l \in [-1,1]\). 
Since we want our controller emulate realistic human motion, we need to derive a
smooth controller where we allow \(l \in [\underbar{l},\overbar{l}]\); then, we
will prove that it injects energy.

Figure \ref{fig:vlp-optimal-controller-qp-plane} displays \(l_\text{opt}\) 
directly on the \((q,p)\)-plane.
Because the length remains constant inside each quadrant and changes only when it
crosses one of the axes, one can use the angle \(\theta = \arctan_2(p,q)\) to
redefine the time-optimal controller in a way which is easier to ``smooth out".
Figure \ref{fig:vlp-optimal-controller} shows the time-optimal pumping strategy
as a function \(l^\star(\theta)\) (\ref{eqn:vlp-optimal-controller}), where now
the length varies between \([\underbar{l},\overbar{l}]\) rather than 
\([-1,1]\).

\begin{equation}\label{eqn:vlp-optimal-controller}
   l^\star(\theta):= \begin{cases}
      \overline{l} & \theta \in [-\frac{\pi}{2},0[ \cup [\frac{\pi}{2}, \pi[ \\
      \underline{l} & \theta \in [-pi, -\frac{\pi}{2}[ \cup [0,\frac{\pi}{2}[ \\
   \end{cases}
\end{equation}

\begin{figure}
   \centering
   \begin{subfigure}[t]{0.45\textwidth}
      \includestandalone[width=\linewidth]{images/vlp_optimal_controller_qp_plane}
      \caption{The time-optimal controller \(l_\text{opt}\) mapped onto the \((q,p)\)-plane.}
      \label{fig:vlp-optimal-controller-qp-plane}
   \end{subfigure}
   \hfill
   \begin{subfigure}[t]{0.45\textwidth}
      \includestandalone[width=\linewidth]{images/vlp_optimal_controller}
      \caption{The time-optimal controller converted to the alternate
         representation \(l^\star(\theta)\).}
      \label{fig:vlp-optimal-controller}
   \end{subfigure}
   \caption{The time-optimal controller for a standing swing as derived by
      \cite{pumping_swing_standing_squatting}. The colour red
      corresponds to standing, blue to squatting, and 
      \(\theta := \arctan_2(p,q)\) is the angle of the VLP phase in the
      \((q,p)\)-plane.}
\end{figure}

We now define a continuous function which
approximates \(l^\star(\theta)\) by attaching sinusoids at the
discontinuities (see Figure \ref{fig:vlp-T-controller}). 
Let \(\Delta l := (\overline{l} - \underline{l})/2\) and 
\(l_{\text{avg}} := (\overline{l} + \underline{l})/2\).
Providing these newly attached sinusoids with a frequency \(\omega = \frac{\pi}{T}\) 
(where \(T \in \, ]0,\frac{\pi}{2}]\) is a control parameter), we get the
continuously-differentiable controller \(l_T(\theta)\) (\ref{eqn:vlp-T-controller}).
\begin{equation}\label{eqn:vlp-T-controller}
   l_T(\theta) = \begin{cases}
      \overline{l} & \theta \in \left[-\frac{\pi}{2} + \frac{T}{2}, -\frac{T}{2}\right] 
      \cup \left[\frac{\pi}{2} + \frac{T}{2}, \pi - \frac{T}{2}\right] \\
      \underline{l} & \theta \in \left[-\pi + \frac{T}{2}, -\frac{\pi}{2} - \frac{T}{2}\right] 
      \cup \left[\frac{T}{2}, \frac{\pi}{2} - \frac{T}{2}\right] \\
      -\Delta l \sin(\omega(\theta + \pi)) + l_{\text{avg}} & \theta \in
      \left[-\pi,-\pi + \frac{T}{2}\right] \\
      -\Delta l \sin(\omega \theta) + l_\text{avg} & \theta \in [-\frac{T}{2},
      \frac{T}{2}] \\
      \Delta l \sin(\omega(\theta - a)) + l_\text{avg} & 
      \theta \in \left[a - \frac{T}{2}, a + \frac{T}{2}\right] \text{ for } 
      a \in \left\{-\frac{\pi}{2}, \frac{\pi}{2}\right\} \\
      -\Delta l \sin(\omega(\theta-\pi)) & \theta \in \left[\pi - \frac{T}{2},\pi\right] \\
   \end{cases}
\end{equation}

\begin{figure}
   \centering
   \includestandalone[width=0.7\textwidth]{images/vlp_T_controller}
   \caption{The continuous VNHC-style VLP controller \(l_T(\theta)\).}\label{fig:vlp-T-controller}
\end{figure}

This controller approximates the optimal controller because 
\[
   \lim\limits_{T \rightarrow 0} l_T(\theta) = l^\star(\theta)
\]
Unfortunately, while \(l_T(\theta)\) is continuously-differentiable, 
it is not twice-differentiable for most values of \(T\).
If we wish to use it as a VNHC, we must ensure that either the generalized
forces \(\tau\) acting on \(p_l\) can be discontinuous (which is certainly
achievable by humans), or we must find a value of \(T\) where this controller is
smooth.
Thankfully, \(l_{\frac{\pi}{2}}(\theta)\) is smooth as it can be simplified into
the expression (\ref{eqn:vlp-smoothed-controller}). 
\begin{equation}\label{eqn:vlp-smoothed-controller}
   l_\frac{\pi}{2}(\theta) = -\Delta l \sin(2\theta) + l_{\text{avg}}
\end{equation}
It is plotted for demonstration in Figure \ref{fig:vlp-smoothed-controller}.

\begin{figure}
   \centering
   \includestandalone[width=0.6\textwidth]{images/vlp_smoothed_controller}
   \caption{The smoothed VLP controller which corresponds to the VNHC \(l =
      l_\frac{\pi}{2}(\theta)\).}\label{fig:vlp-smoothed-controller}
\end{figure}

Because \(l_\frac{\pi}{2}(\theta)\) is smooth and it approximates
\(l^\star(\theta)\), we set our VNHC to be
\[
   h(q,p) = l - l_\frac{\pi}{2}\left(\theta(q,p)\right)
\]
We still need to prove this VNHC will inject energy into a VLP. 
Before we can do this, we require the following lemma.

>>>>

The motion of the VLP has been well studied (see for instance
\cite{dynamics_periodic_vlp}), and many control mechanisms exist
to stabilize trajectories of the system. While many of these controllers 
are time-dependent, \citet{vlp_energy_shaping}
offer a time-independent technique to inject energy into the VLP. 
They design a controller through a technique called \textit{energy shaping}
and prove that it stabilizes any desired energy level set.
However, their control input depends on a pre-specified target energy and requires
knowledge of the current total energy of the VLP.
This makes their energy injection mechanism ``ad-hoc" in the sense that it is
tailored very specifically to the VLP, and is not generalizable to a larger 
methodology.

It may be better to base the control design on natural biological behaviour.
In this chapter we will make use of the general VNHC framework developed in Chapter
\ref{ch:vnhcs} to add and remove energy from the VLP in a time-independent
manner.
We'll show that, unlike energy shaping, VNHCs can be used to stabilize energy
levels while maintaining the structured motion of a human on a swing.

\section{Dynamics of the Variable Length Pendulum}
% Derive the dynamics of the VLP in Hamiltonian form
We will model the VLP as a point mass \(m\)
connected to a fixed pivot by a massless rod of varying length \(l\) with angle 
\(q \in \mathbb{S}^1\) from the vertical, as in Figure
\ref{fig:vlp-model}. 
We will ignore any damping and frictional forces in this model.
In a realistic VLP, the rod length \(l\) varies between some minimum
length \(\underline{l} \geq 0\) and some maximum length 
\(\overline{l} > \underline{l}\). The configuration of the VLP is the vector
\(\mathbf{q} := (q,l) \in \mathbb{S}^1 \times [\underline{l},\overline{l}]\).

@@

Using this configuration, we will compute the Hamiltonian dynamics of the system.
The cartesian position of the mass at the tip of the pendulum
is given by \(x = (l\sin(q),-l\cos(q))\), while its velocity is
\(\dot{x} = (\dot{l}\sin(q) + l\cos(q)\dot{q}, -\dot{l}\cos(q) + l\sin(q)\dot{q})\).
Computing the kinetic energy \(T\) yields
\[
   T(\mathbf{q},\dot{\mathbf{q}}) = 
   \frac{1}{2}m\norm{\dot{x}}^2 = \frac{1}{2}m\left(\dot{l}^2 + l^2\dot{q}^2\right)
   .
\]
The potential energy \(P\) with respect to the pivot (under a gravitational
acceleration \(g\)) is
\[
   P(\mathbf{q}) = -mgl\cos(q)
   .
\]
Collecting the kinetic energy into a quadratic form, we get the Lagrangian
\[
   \mathcal{L}(\mathbf{q},\dot{\mathbf{q}}) 
   = \frac{1}{2} \dot{\mathbf{q}}\tpose D(\mathbf{q})\dot{\mathbf{q}} - P(\mathbf{q})
   = \frac{1}{2}
   \begin{bmatrix} \dot{q} & \dot{l} \end{bmatrix}
   \begin{bmatrix}
      ml^2 & 0 \\
      0 & m \\
   \end{bmatrix}
   \begin{bmatrix} 
      \dot{q} \\ \dot{l}
   \end{bmatrix}
   + mgl\cos(q)
   .
\]
Computing the conjugate of momenta to \(\mathbf{q}\), we get 
\[
   \mathbf{p} := \begin{bmatrix} p \\ p_l \end{bmatrix} 
   = \begin{bmatrix} ml^2\dot{q} \\ m\dot{l} \end{bmatrix} 
   .
\]
Performing the Legendre transform on \(\mathcal{L}\) and setting
\(M(\mathbf{q}) := D(\mathbf{q})\), \(V(\mathbf{q}) := P(\mathbf{q})\),
we get the Hamiltonian (\ref{eqn:simply-actuated-hamiltonian}) whose
dynamics (\ref{eqn:simply-actuated-dynamics}) resolve to  
\begin{align}\label{eqn:vlp-hamiltonian-with-pl}
   \mathcal{H} &= \frac{1}{2} \begin{bmatrix} p & p_l \end{bmatrix}
      \begin{bmatrix}
         \frac{1}{ml^2}  & 0 \\
         0 & \frac{1}{m}
      \end{bmatrix} \begin{bmatrix} p \\ p_l \end{bmatrix} - mgl\cos(q)
      , \\
     &\begin{cases}
        \dot{q} = \frac{p}{ml^2} \\
        \dot{l} = \frac{p_l}{m} \\
        \dot{p} = -mgl\sin(q) \\
        \dot{p}_l = \frac{p^2}{ml^3} + mg\cos(q) + \tau
        . \\
   \end{cases} \nonumber
\end{align}
The control input is a force \(\tau \in \mathbb{R}\) affecting the dynamics of
\(p_l\), acting colinearly with the rod.
We assume the force does not affect the dynamics of \(p\) in any way -
that is, the control input cannot enact any lateral force on the pendulum.
This is extremely useful because it means the VLP is described in simply
actuated coordinates, which allows us to apply the theory of VNHCs we
developed in Chapter \ref{ch:vnhcs}.

Let us define the VNHC \(l = L(q,p)\) -- by which we mean we are
actually defining the VNHC \(h(q,p) = l - L(q,p) = 0\).
Assuming this is a regular VNHC, the constraint manifold is
\(\Gamma = \mathbb{S}^1 \times \R\), which is parameterized by the unactuated
phase \((q,p)\).
The constrained dynamics are Hamiltonian, and Theorem \ref{thm:zero-dynamics}
tells us they are described entirely by \((\dot{q},\dot{p})\) with \(l\)
replaced by \(L(q,p)\):
\begin{align}\label{eqn:vlp-hamiltonian}
   \mathcal{H}(q,p) &= \frac{1}{2} \frac{p^2}{mL^2} - \frac{1}{2}\dot{L}^2 -
      mgL\cos(q)
   , \\
  &\begin{cases}
      \dot{q} = \frac{p}{m L^2} \\
      \dot{p} = -mgL\sin(q)
      . \\ 
   \end{cases} \nonumber
\end{align}
Note that we suppress the function notation for \(L(q,p)\) and
\(\dot{L}(q,p)\) for clarity.

The total mechanical energy of the system restricted to the constraint manifold
is given by (\ref{eqn:vlp-energy}), where \(L(q,p)\) and \(\dot{L}(q,p)\) are
known.
\begin{equation}\label{eqn:vlp-energy}
   E(q,p) = \frac{1}{2} \frac{p^2}{mL^2} + \frac{1}{2}\dot{L}^2 - mgL\cos(q)
\end{equation}
In the rest of this chapter, we will derive a \(C^2\) function \(L(q,p)\) 
which produces constrained dynamics that inject energy into the VLP in a
human-like manner.

@@

\section{The VLP Constraint}
% Go through the development of the constraint and prove it gains energy
To motivate why a VNHC could inject energy into the VLP in a human-like
manner, we will examine a person standing on a swing.
As can be seen in Figure \ref{fig:child-vlp}, a person's center of mass moves closer
to the swing's pivot when they stand, and moves away from the pivot when they
squat.
This is equivalent to the VLP model from Figure \ref{fig:vlp-model}, where
standing and squatting correspond to shortening and lengthening the pendulum
(respectively).
\begin{figure}
   \centering
   \begin{subfigure}[t]{0.45\textwidth}
      \includegraphics[]{images/child_vlp_standing.png}
      \caption{A person standing on a swing has their center of mass 
      close to the pivot.}
   \end{subfigure}
   \hfill
   \begin{subfigure}[t]{0.45\textwidth}
      \includegraphics[]{images/child_vlp_squatting.png}
      \caption{When a person squats on a swing, their center of mass extends
      away from the pivot.}
   \end{subfigure}
   \caption{The VLP representation of a person on a standing swing.}
   \label{fig:child-vlp}
\end{figure}

@@

Notice that the child's pumping strategy requires knowledge of
when the system is at the ``bottom" or ``top" of the swing. 
Since being at the ``bottom" is equivalent to having angle \(q = 0\) and being
at the ``top" is equivalent to having momentum \(p = 0\), a controller based on
this strategy will necessarily use the full phase \((q,p)\). 
This is why we must use VNHCs instead of VHCs to perform this maneuver.

The time-optimal controller from \cite{pumping_swing_standing_squatting} is (in
our notation)
\[
   L^\star(q,p) := -\sign{qp}
   ,
\]
which is a piecewise-continuous controller that varies between \(\pm 1\). 
We could set our constraint to be \(l = L^\star(q,p)\), but this is not a VNHC
because it is not \(C^2\).
It also makes the assumption that \(l \in [-1,1]\) and can switch
instantaneously.
Since need to enforce the constraint using the physical input \(\tau\) (which
ideally would emulate realistic human motion), we cannot use \(L^\star\) as our
VNHC.
We will instead find an alternate representation of \(L^\star\) which can be
converted into a VNHC, and which allows 
\(l \in [\underbar{l},\overbar{l}]\).

Figure \ref{fig:vlp-optimal-controller-qp-plane} displays \(L^\star(q,p)\) 
on the \((q,p)\)-plane.
Because the length remains constant inside each quadrant and changes only when it
crosses one of the axes, one can use the angle \(\theta = \arctan_2(p,q)\) to
redefine the time-optimal pumping strategy as a function of \(\theta\).
Abusing notation, we denote this by \(L^\star(\theta)\), which is defined in
(\ref{eqn:vlp-optimal-controller}).
Figure \ref{fig:vlp-optimal-controller} shows the graph of \(L^\star(\theta)\), 
where now the length varies between \([\underbar{l},\overbar{l}]\) rather than 
\([-1,1]\).

\begin{equation}\label{eqn:vlp-optimal-controller}
   L^\star(\theta):= \begin{cases}
      \overline{l} & \theta \in [-\frac{\pi}{2},0[ \cup [\frac{\pi}{2}, \pi[ \\
      \underline{l} & \theta \in [-pi, -\frac{\pi}{2}[ \cup [0,\frac{\pi}{2}[ \\
   \end{cases}
   .
\end{equation}

\begin{figure}
   \centering
   \begin{subfigure}[t]{0.45\textwidth}
      \includestandalone[width=\linewidth]{images/vlp_optimal_controller_qp_plane}
      \caption{The time-optimal controller \(L^\star(q,p)\) mapped onto the
      \((q,p)\)-plane. Here, red corresponds to \(L^\star(q,p) = -1\) and blue
      to \(L^\star(q,p) = 1\).}
      \label{fig:vlp-optimal-controller-qp-plane}
   \end{subfigure}
   \hfill
   \begin{subfigure}[t]{0.45\textwidth}
      \includestandalone[width=\linewidth]{images/vlp_optimal_controller}
      \caption{The time-optimal controller converted to the alternate
      representation \(L^\star(\theta)\).}
      \label{fig:vlp-optimal-controller}
   \end{subfigure}
   \caption{The time-optimal controller for a standing swing as derived by
      \cite{pumping_swing_standing_squatting}. The colour red
      corresponds to standing, blue to squatting, and 
      \(\theta := \arctan_2(p,q)\) is the angle of the VLP phase in the
      \((q,p)\)-plane.}
\end{figure}

We now define a continuous function which
approximates \(L^\star(\theta)\) by attaching sinusoids at the
discontinuities (see Figure \ref{fig:vlp-T-controller}). 
Let \(\Delta l := (\overline{l} - \underline{l})/2\) and 
\(l_{\text{avg}} := (\overline{l} + \underline{l})/2\).
Providing these newly attached sinusoids with a frequency \(\omega = \frac{\pi}{T}\) 
(where \(T \in \, ]0,\frac{\pi}{2}]\) is a control parameter), we get the
continuously-differentiable controller \(L_T(\theta)\) (\ref{eqn:vlp-T-controller}).
\begin{equation}\label{eqn:vlp-T-controller}
   L_T(\theta) = \begin{cases}
      \overline{l} & \theta \in \left[-\frac{\pi}{2} + \frac{T}{2}, -\frac{T}{2}\right] 
      \cup \left[\frac{\pi}{2} + \frac{T}{2}, \pi - \frac{T}{2}\right] \\
      \underline{l} & \theta \in \left[-\pi + \frac{T}{2}, -\frac{\pi}{2} - \frac{T}{2}\right] 
      \cup \left[\frac{T}{2}, \frac{\pi}{2} - \frac{T}{2}\right] \\
      -\Delta l \sin(\omega(\theta + \pi)) + l_{\text{avg}} & \theta \in
      \left[-\pi,-\pi + \frac{T}{2}\right] \\
      -\Delta l \sin(\omega \theta) + l_\text{avg} & \theta \in [-\frac{T}{2},
      \frac{T}{2}] \\
      \Delta l \sin(\omega(\theta - a)) + l_\text{avg} & 
      \theta \in \left[a - \frac{T}{2}, a + \frac{T}{2}\right] \text{ for } 
      a \in \left\{-\frac{\pi}{2}, \frac{\pi}{2}\right\} \\
      -\Delta l \sin(\omega(\theta-\pi)) & \theta \in \left[\pi - \frac{T}{2},\pi\right] \\
   \end{cases}
   .
\end{equation}

\begin{figure}
   \centering
   \includestandalone[width=0.7\textwidth]{images/vlp_T_controller}
   \caption{The continuous VLP constraint \(L_T(\theta)\).}\label{fig:vlp-T-controller}
\end{figure}

This family of constraints approximates \(L^\star(\theta)\) because 
\[
   \lim\limits_{T \rightarrow 0} L_T(\theta) = L^\star(\theta)
   .
\]
Unfortunately, while \(L_T(\theta)\) is continuously-differentiable, 
it is not twice-differentiable for most values of \(T\).
If we wish to use it as a VNHC, we must ensure that either the generalized
forces \(\tau\) acting on \(p_l\) can be discontinuous (which is certainly
achievable by humans), or we must find a value of \(T\) where this constraint is
at least \(C^2\).
Thankfully, setting \(T = \frac{\pi}{2}\) yields the smooth function 
\(L_{\frac{\pi}{2}}(\theta)\), which can be simplified into
\begin{equation}\label{eqn:vlp-smoothed-controller}
   L_\frac{\pi}{2}(\theta) = -\Delta l \sin(2\theta) + l_{\text{avg}}
   .
\end{equation}
This smooth constraint is plotted for demonstration in Figure
\ref{fig:vlp-smoothed-controller}.

\begin{figure}
   \centering
   \includestandalone[width=0.6\textwidth]{images/vlp_smoothed_controller}
   \caption{The smoothed VLP constraint which corresponds to the VNHC 
      \(l = L_\frac{\pi}{2}(\theta)\).}\label{fig:vlp-smoothed-controller}
\end{figure}

Because \(L_\frac{\pi}{2}(\theta)\) is smooth and it approximates
\(L^\star(\theta)\), we set our VNHC to be
\[
   h(q,p) = l - L_\frac{\pi}{2}\left(\theta(q,p)\right)
   .
\]
We still need to prove this will inject energy into the VLP. 
Before we can do this, we require the following lemma.
